---
title: Crossfit Open 2019 --- an Analysis
author: Stefan Heyder
date: March 26, 2019
output: html_document
---
```{r}
library(tidyverse)
```


# The Crossfit Open

# Data
```{r}
crossfit.open <- read_csv("clean_data/2019-03-26_16:21:24_clean_data.csv") %>%
    mutate(WOD = as.character(WOD)) %>%
    filter(Repetitions > 0) # Remove scaled
```

# 19.1


This workout is easy to analyze, as there is only one metric determining your score: How many repetitions of wall balls and rowing did you accomplish in the given time cap? 
We will thus start our analysis with a density plot of the repetition for 19.1:

```{r}
(plot.19.1 <- crossfit.open %>%
    filter(WOD == '19.1') %>%
ggplot(aes(x = Repetitions)) + geom_density()
)
```

There seem to be many modes for this distribution. 
One possible reason could be the different rounds, as athletes may push at the end of a workout to "finish the round" they are currently in. 
One round consists of 19 wall-ball shots and 19 kcal on the rower, thus we'd expect a peak every 38 Repetitions.
I'll also add some lines for every half round, after the athlete has finished their set of 19 wall-balls.
Let's add indicators for these rounds to our plot

```{r}
rounds <- rbind(tibble(reps = 38 * 1:10, Rounds = "Full"), tibble(reps =  19 + 38 * 1:10, Rounds = "Half"))
plot.19.1 +
    geom_vline(data = rounds, aes(xintercept = reps, color = Rounds), linetype = 2, alpha = 0.4)

```




